---
title: "Change over time in different currencies"
output: 
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = F, warning = F, message = F)
library(BBSsize)
library(rwar)
library(ggplot2)
library(dplyr)
theme_set(theme_bw())
```

# One route

We will work with the BBS route for Granby, CT to demonstrate the workflow that gets applied to all the routes. (Just RMD's usual guinea pig route). 

## Pull and examine raw data

```{r, fig.dim = c(8, 3)}

dat <- granby

dat_isd <- just_isd(granby)

dat_to_plot <- dat_isd %>%
  mutate(energy = estimate_b(mass)) %>%
  filter(year %in% c(1988:1992, 2014:2018)) %>%
  group_by_all() %>%
  mutate(timeperiod = ifelse(year > 2000, "end", "begin")) %>%
  ungroup() %>%
  group_by(year) %>%
  summarize(
    total_abundance = dplyr::n(),
    total_energy = sum(energy),
    total_biomass = sum(mass)
  ) %>%
  ungroup()

gridExtra::grid.arrange(grobs = list(ggplot(dat_to_plot, aes(year, total_abundance)) + geom_point() + ggtitle("abundance"),
ggplot(dat_to_plot, aes(year, total_energy)) + geom_point() + ggtitle("energy"),
ggplot(dat_to_plot, aes(year, total_biomass)) + geom_point() + ggtitle("biomass")), ncol = 3)


```


Granby looks like not an appreciable change in abundance, but potentially an increase in energy and more probably biomass.

## Test for size-structure-driven decoupling

We want to test whether the changes in energy and biomass exceed what we would expect to see if there had been **no change** in the size structure from the begin to the end period.


From here on out, code is taken directly from `View::rwar::draw_communities` and narrated.

### Arguments and explanation

`draw_communities(ts_comp, begin_years = 1988:1992, end_years = 2014:2018, draw_seed = NULL, sampling_gmms = NULL, initial_isd_seed = NULL, raw_isd_seed = NULL)`

```{r}
ts_comp = dat
begin_years = 1988:1992
end_years = 2014:2018
draw_seed = NULL
sampling_gmms = NULL
inital_isd_seed = NULL
raw_isd_seed = NULL

```

### Construct sampling gmms

"Sampling GMMS" are density functions from Gaussian mixture models that have a probability density for every mass along the spectrum of possible masses (and then a large buffer). Drawing from `$mass` with probability `$density` is a random number generator for sizes. 

```{r}

  # If GMM density smooths to sample from are not provided, get them
  # When running at scale, you want to provide them: 1) to ensure that all sims are being drawn from the same sampling GMMS, 2) for speed, it takes a long time to fit a GMM to a lot of ISD draws.

  if(is.null(sampling_gmms)) {

    sampling_gmms <- construct_sampling_gmm(ts_comp, begin_years = begin_years, end_years = end_years, initial_isd_seed = NULL)

  }
```

```{r}
ggplot(sampling_gmms$begin, aes(mass, density)) + geom_line() + geom_line(data = sampling_gmms$end, color = "green") + ggtitle("Sampling GMMs", subtitle = "Black = begin, green = end") + xlab("Mass (log)")
```

### Get template ISD dataframes

```{r}

  # This doesn't actually matter, I am using the raw isds to get the correct shape data frame. I am keeping it this way, even though it's inefficient, because in the back of my mind I think I might want to pull out raw estimates for biomass and energy use (i.e. drawing individuals directly from normal distributions for each year, rather than drawing from the ISD combining over all years). It'll be a tiny change but I'm not 100% sure what I want to do with that info yet so holding off.

  if(is.null(raw_isd_seed)) {
    set.seed(NULL)
    raw_isd_seed <- sample(1:1000000, 1)
  }

  # Here I am sampling ISDs to get dfs of the correct shape to then sample new body masses from different density fxns.

  raw_isd <- BBSsize::simulate_isd_ts(ts_comp, isd_seed = raw_isd_seed)$isd

  begin_isd <- dplyr::filter(raw_isd, year %in% begin_years) %>%
    dplyr::mutate(timeperiod = "begin")
  end_isd <- dplyr::filter(raw_isd, year %in% end_years) %>%
    dplyr::mutate(timeperiod = "end")
  
```


### Draw individuals corresponding to "sim" (no size change) and actual dynamics

```{r}

  # Draw individuals for each time period from the MATCHING density functions
  # This will destroy interannual, intratimeperiod variation in the size structure, which we're OK with (the point of using 5-year intervals is to smooth out species accumulation)
  # I do not think it is a good idea to provide draw_seed, that will constrain things to come out the same in weird ways at this scale.
  begin_individuals <- add_drawn_individuals(begin_isd, sampling_gmms$begin, draw_seed = draw_seed)

  end_individuals <- add_drawn_individuals(end_isd, sampling_gmms$end, draw_seed = draw_seed)

  actual_individuals <- dplyr::bind_rows(begin_individuals, end_individuals) %>%
    dplyr::mutate(source = "actual")


  # Now draw individuals for each time period with scrambled ISDs. Specifically, draw for the beginning from the beginning ISD. But then also draw the end from the beginning ISD. This gives an "end" ISD pretending that the ISD didn't change from the beginning.
  begin_individuals_sim <- add_drawn_individuals(begin_isd, sampling_gmms$begin, draw_seed = draw_seed)

  end_individuals_sim <- add_drawn_individuals(end_isd, sampling_gmms$begin, draw_seed = draw_seed)


  sim_individuals <- dplyr::bind_rows(begin_individuals_sim, end_individuals_sim) %>%
    dplyr::mutate(source = "sim")

```


### Combine with "raw"

The "raw" ISDs have not been in-and-out of the GMM smooth and then resampling. The resampling is necessary to make the nullmodel/sims here possible, but it masks intraannual variability in the size spectrum and it can downweight (very slightly) the density of large individuals.

```{r}

  # Go ahead and pull the raw state variable estimates too..

  raw_individuals <- dplyr::bind_rows(begin_isd, end_isd) %>%
    dplyr::mutate(energy = BBSsize::estimate_b(mass),
                  source = "raw",
                  isd_timeperiod = "raw",
                  sampling_seed = NA)

  all_individuals <- dplyr::bind_rows(actual_individuals, sim_individuals, raw_individuals)


```

### Compute time-period level state variables

```{r}


  # Summarize individuals to get toal abundance, biomass, and energy use per year for each sim scenario.
  # And add route-level identifying info.
  all_svs <- all_individuals %>%
    dplyr::group_by(year, timeperiod, isd_timeperiod, sampling_seed, isd_seed, source) %>%
    dplyr::summarize(total_abundance = dplyr::n(),
                     total_biomass = sum(mass),
                     total_energy = sum(energy)) %>%
    dplyr::ungroup() %>%
    dplyr::bind_cols(as.data.frame(ts_comp$metadata$location)) %>%
    dplyr::mutate(matssname = paste0("bbs_rtrg_", route, "_", statenum))


```
