---
title: "Portal EE v CC"
output: 
    github_document:
        toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(BBSsize)
library(portalr)
library(dplyr)
library(ggplot2)
library(drake)
theme_set(theme_bw())
```

# Thinking

Trying to use Portal exclosure v. controls to develop intuition for how this null model works.

The thinking is that:

- EE v CC in the early 1980s should constitute a pretty pronounced size shift, and so ISD similarity should be on the low end of what is expected given species turnover. This is because the primary difference between the exclosures and controls is the removal of the 3 largest species. 

- Contrasting with EE v CC in the early 2000s, when PB was dominant on EE plots. At that time, the biggest difference - from our perspective - is that DM/DO are in that size class on CC, but PB is dominating that size class on EE plots. Which I would think constitutes pretty good size-structured replacement (although note that PB is slightly smaller than the krats). So the thinking is that that comparison would be on the **high** end of ISD similarity | species turnover. 

- However, either of these could be not-what-happens because of
    - other changes along the size spectrum (idk, reithros or something)
    - the ISD being inherently kind of....squishy and conserved. 
    - species pool effects - haven't thought these through completely
    - this null model seems to give prety broad 95% intervals/it seems pretty hard to deviate from this null. Part of the purpose of this is to gauge if - even in scenarios that **I** think are Definitely Size Structured - deviations don't show up. 


# Early 1980s

```{r}


## Set up the cache and config
db <- DBI::dbConnect(RSQLite::SQLite(), here::here("drake-cache-pn.sqlite"), synchronous = NULL)
cache <- storr::storr_dbi("datatable", "keystable", db)

cache$del(key = "lock", namespace = "session")

loadd(siteDescs_exclosure_real_counts, cache = cache)
loadd(siteDescs_control_real_counts, cache = cache)
loadd(siteDescs_control_shuffled_counts_sim_splist_2L, cache = cache)
loadd(siteDescs_exclosure_shuffled_counts_sim_splist_2L, cache = cache)
loadd(allComps, cache = cache)

```


```{r}

eighties_isd_gmms <- list(control = siteDescs_control_real_counts$gmm, exclosure = siteDescs_exclosure_real_counts$gmm) %>%
  bind_rows(.id = "trt")
ggplot(eighties_isd_gmms, aes(mass, density, color = trt)) + 
  geom_line()
# 
# eighties_isd_sim_gmms <- list(control = siteDescs_control_shuffled_counts_sim_splist_2L$gmm, exclosure = siteDescs_exclosure_shuffled_counts_sim_splist_2L$gmm) %>%
#   bind_rows(.id = "trt")
# ggplot(eighties_isd_sim_gmms, aes(mass, density, color = trt)) + 
#   geom_line() 

```

```{r}

ggplot(filter(allComps, sim > 0), aes(isd_overlap)) +
  geom_histogram() +
  geom_point(data = filter(allComps, sim < 0), aes(isd_overlap, 0), color = "red", size = 5)

simComps <- filter(allComps, sim > 0)
realComp <- filter(allComps, sim < 0)

# percentile

sum(simComps$isd_overlap > realComp$isd_overlap[1]) / nrow(simComps)

# ses 

(realComp$isd_overlap[1] - mean(simComps$isd_overlap)) / sd(simComps$isd_overlap)

#filter(simComps, isd_overlap < realComp$isd_overlap[1])

realComp
```




```{r}

DBI::dbDisconnect(db)
rm(cache)
rm(list =ls())
```

# Early 2000s


```{r}


## Set up the cache and config
db <- DBI::dbConnect(RSQLite::SQLite(), here::here("drake-cache-pn-t.sqlite"), synchronous = NULL)
cache <- storr::storr_dbi("datatable", "keystable", db)

cache$del(key = "lock", namespace = "session")

loadd(siteDescs_exclosure_real_counts, cache = cache)
loadd(siteDescs_control_real_counts, cache = cache)
loadd(siteDescs_control_shuffled_counts_sim_splist_2L, cache = cache)
loadd(siteDescs_exclosure_shuffled_counts_sim_splist_2L, cache = cache)
loadd(allComps, cache = cache)

```


```{r}

thous_isd_gmms <- list(control = siteDescs_control_real_counts$gmm, exclosure = siteDescs_exclosure_real_counts$gmm) %>%
  bind_rows(.id = "trt")
ggplot(thous_isd_gmms, aes(mass, density, color = trt)) + 
  geom_line()
# 
# thous_isd_sim_gmms <- list(control = siteDescs_control_shuffled_counts_sim_splist_2L$gmm, exclosure = siteDescs_exclosure_shuffled_counts_sim_splist_2L$gmm) %>%
#   bind_rows(.id = "trt")
# ggplot(thous_isd_sim_gmms, aes(mass, density, color = trt)) + 
#   geom_line() 

```

```{r}

ggplot(filter(allComps, sim > 0), aes(isd_overlap)) +
  geom_histogram() +
  geom_point(data = filter(allComps, sim < 0), aes(isd_overlap, 0), color = "red", size = 5)

simComps <- filter(allComps, sim > 0)
realComp <- filter(allComps, sim < 0)

# percentile

sum(simComps$isd_overlap < realComp$isd_overlap[1]) / nrow(simComps)

# ses 

(realComp$isd_overlap[1] - mean(simComps$isd_overlap)) / sd(simComps$isd_overlap)

realComp

```

```{r}

DBI::dbDisconnect(db)
rm(cache)
rm(list =ls())
```

# Early 2000s - manipulated


```{r}


## Set up the cache and config
db <- DBI::dbConnect(RSQLite::SQLite(), here::here("drake-cache-pn-t-M.sqlite"), synchronous = NULL)
cache <- storr::storr_dbi("datatable", "keystable", db)

cache$del(key = "lock", namespace = "session")

loadd(siteDescs_exclosure_real_counts, cache = cache)
loadd(siteDescs_control_real_counts, cache = cache)
loadd(siteDescs_control_shuffled_counts_sim_splist_2L, cache = cache)
loadd(siteDescs_exclosure_shuffled_counts_sim_splist_2L, cache = cache)
loadd(allComps, cache = cache)

```


```{r}

thous_isd_gmms <- list(control = siteDescs_control_real_counts$gmm, exclosure = siteDescs_exclosure_real_counts$gmm) %>%
  bind_rows(.id = "trt")
ggplot(thous_isd_gmms, aes(mass, density, color = trt)) + 
  geom_line()

thous_isd_sim_gmms <- list(control = siteDescs_control_shuffled_counts_sim_splist_2L$gmm, exclosure = siteDescs_exclosure_shuffled_counts_sim_splist_2L$gmm) %>%
  bind_rows(.id = "trt")
ggplot(thous_isd_sim_gmms, aes(mass, density, color = trt)) + 
  geom_line() 

```

```{r}

ggplot(filter(allComps, sim > 0), aes(isd_overlap)) +
  geom_histogram() +
  geom_point(data = filter(allComps, sim < 0), aes(isd_overlap, 0), color = "red", size = 5)

simComps <- filter(allComps, sim > 0)
realComp <- filter(allComps, sim < 0)

# percentile

sum(simComps$isd_overlap < realComp$isd_overlap[1]) / nrow(simComps)

# ses 

(realComp$isd_overlap[1] - mean(simComps$isd_overlap)) / sd(simComps$isd_overlap)
realComp


```

# Over time


```{r}


DBI::dbDisconnect(db)
rm(cache)
rm(list =ls())
## Set up the cache and config
db <- DBI::dbConnect(RSQLite::SQLite(), here::here("drake-cache-pn-w.sqlite"), synchronous = NULL)
cache <- storr::storr_dbi("datatable", "keystable", db)

cache$del(key = "lock", namespace = "session")

loadd(siteDescs_start_real_counts, cache = cache)
loadd(siteDescs_end_real_counts, cache = cache)
loadd(siteDescs_end_shuffled_counts_sim_splist_2L, cache = cache)
loadd(siteDescs_start_shuffled_counts_sim_splist_2L, cache = cache)
loadd(allComps, cache = cache)

```


```{r}

thous_isd_gmms <- list(end = siteDescs_end_real_counts$gmm, start = siteDescs_start_real_counts$gmm) %>%
  bind_rows(.id = "trt")
ggplot(thous_isd_gmms, aes(mass, density, color = trt)) + 
  geom_line()
# 
# thous_isd_sim_gmms <- list(control = siteDescs_control_shuffled_counts_sim_splist_2L$gmm, exclosure = siteDescs_exclosure_shuffled_counts_sim_splist_2L$gmm) %>%
#   bind_rows(.id = "trt")
# ggplot(thous_isd_sim_gmms, aes(mass, density, color = trt)) + 
#   geom_line() 

```

```{r}

ggplot(filter(allComps, sim > 0), aes(isd_overlap)) +
  geom_histogram() +
  geom_point(data = filter(allComps, sim < 0), aes(isd_overlap, 0), color = "red", size = 5)

simComps <- filter(allComps, sim > 0)
realComp <- filter(allComps, sim < 0)

# percentile

sum(simComps$isd_overlap < realComp$isd_overlap[1]) / nrow(simComps)

# ses 

(realComp$isd_overlap[1] - mean(simComps$isd_overlap)) / sd(simComps$isd_overlap)


realComp

```

```{r}

DBI::dbDisconnect(db)
rm(cache)
rm(list =ls())
```
