---
title: "Modeling approach & validation"
output: 
        github_document:
            toc: TRUE
            df_print: kable
author: Renata Diaz
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(BBSsize)
library(ggplot2)
library(dplyr)
theme_set(theme_bw())
```

# How we get here

See 02_01_change_over_time_sims.Rmd for an explanation of how we generate sims describing energy use/biomass dynamics assuming a) actual change in the size structure over time or b) no change in the size structure (such that any change that occurs is driven solely by changes in abundance).

From there, we would like to know:

- how much change has occurred due to change in abundance
- how much change in the currency deviates from the change driven by change in abundance (reflecting change in the size structure)


# Data 

We use the sims method to get estimates of total energy or total biomass for each year under the scenarios 1) actual size structure change 2) no change, achieved by using the size structure for the first 5 years for the last 5 years as well. 

```{r, message=F, warning=F,results=F}
#dat = granby

#actual_sims <- rwar::make_actual_sims(dat) # this is a wrapper fxns that draws 100 sims

print(rwar::make_actual_sims)

```

This gives us a dataframe with columns:

```{r}

#colnames(actual_sims)

```

```{r}

#ggplot(actual_sims, aes(year, total_biomass, color = source)) + geom_point() + geom_smooth(method = "lm", se = F)

```

We want to know:

- the slope for abundance
- the slope for currency
- i.e. whether the slope for currency is different from the slope for abundance

To fit models, we need to summarize across all the sims, because including all the individual sims is going to artificially inflate the model's confidence. 

We also want to remove "raw" because that is just for a gut check.

```{r}

#ssims_actual_sims <- rwar::summarize_sims(actual_sims)

print(rwar::summarize_sims)


#head(ssims_actual_sims)

```

# Models

We will use bayesian linear models to estimate the slopes for abundance and currency and the difference between them.

We will compare models of the following forms:

`response ~ timeperiod * source`: Estimates a slope for timeperiod ("begin" or "end") and the interaction between source ("abundance" or "currency")

`response ~ timeperiod`: Only a timeperiod slope, no difference between sources

`response ~ 1`: Only an intercept, no change begin or end.

We will use the LOO criterion from `rstanarm::loo`. Because sometimes the best model according to LOO wins by a very small margin, we will select the simplest model within 1 standard error of the best model for interpretation. 

## Fit models

```{r}

print(rwar::fit_stanlm)

```

## Compare models

```{r}

#fits_compare <- rwar::compare_both_brms(fits)

print(rwar::compare_both_stanarms)

print(rwar::compare_stanarms)

```

## Select best model

```{r}

#fits_winners <- rwar::loo_select(fits_compare)

print(rwar::loo_select)

#print(fits_winners)
```

# Demo scenarios

I ran this pipeline on 4 datasets demonstrating different types of dynamics, and compared the "actual" dynamics to additional sims illustrating different "real" effects.

```{r}
library(drake)

## Set up the cache and config
db <- DBI::dbConnect(RSQLite::SQLite(), here::here("aspirational_structure", "drake_caches", "sim_means-cache-unlocked.sqlite"))
cache <- storr::storr_dbi("datatable", "keystable", db)
cache$del(key = "lock", namespace = "session")

#cached(cache = cache)
loadd(all_sims, cache=cache)

all_sims <- all_sims %>%
  filter(matssname %in% c("bbs_rtrg_116_18", "bbs_rtrg_19_7", "bbs_rtrg_224_3", "bbs_rtrg_318_3"))

```

```{r, fig.dim = c(9,9)}
ggplot(all_sims, aes(year, total_biomass, color = source)) + geom_point() + geom_smooth(method = "lm", se = F, size = .3) + facet_wrap(vars(matssname, simtype), scales = "free", ncol = 3) + theme(legend.position = "bottom")

```

The "sims" correspond to: no change in abunance or size strucutre (no change), no change in size structure (nosizechange). While we expect to get decoupling of currency and size structure in some actual dynamics, we do not expected it in these sims. We have weaker expectations about the slopes of the actual dynamics. Note that the lines are lm smooths without se and not suitable for inference. 

Here are the model winners for the different sims:


224 and 116 both have deviations; 19_7 selects no change; 318_3 selects a slope but not a deviation. None of the sims select overly-complex models. Note that usually but not often we're invoking the one se rule (if `rank` isn't 1 it means we selected a lower-ranked model because it was simpler and within 1 se.)

```{r}
loadd(all_winners, cache = cache)

all_winners %>%
  filter(matssname %in% c("bbs_rtrg_116_18", "bbs_rtrg_19_7", "bbs_rtrg_224_3", "bbs_rtrg_318_3")) %>%
  dplyr::filter(currency == "biomass")
```
