---
title: "5 sites"
output:
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(drake)
library(dplyr)
library(ggplot2)
library(ghibli)

theme_set(theme_bw())
## Set up the cache and config
db <- DBI::dbConnect(RSQLite::SQLite(), here::here("drake-cache-betadiv.sqlite"))
cache <- storr::storr_dbi("datatable", "keystable", db)
cache$del(key = "lock", namespace = "session")

loadd(all_smooths, cache = cache)  
loadd(all_spcomposition, cache = cache)

DBI::dbDisconnect(db)
rm(cache)
rm(db)

```

This is gonna be hacky af, but calculating

- species overlap
- ISD overlap
- **within** each time chunk
- for all possible pairs



```{r}


pull_species_comp <- function(all_spcomp, rt, rg, bcr, chunk) {
  
  return(all_spcomp %>%
           filter(route == rt,
                  region == rg,
                  location.bcr == bcr,
                  timechunk == chunk))
  
}

pull_smooths <- function(all_smooths, rt, rg, bcr, chunk) {
  
  if(chunk == "start") {
    out <- all_smooths %>%
      filter(route == rt,
             region == rg,
             location.bcr == bcr) %>%
      select(route, region, location.bcr, location.statenum, location.routename, mass, start) %>%
      mutate(timechunk = chunk) %>%
      rename(density = start)
  } else {
    out <- all_smooths %>%
      filter(route == rt,
             region == rg,
             location.bcr == bcr) %>%
      select(route, region, location.bcr, location.statenum, location.routename, mass, end) %>%
      mutate(timechunk = chunk) %>%
      rename(density = end)
  }
  
}
compare_species <- function(spcomp1, spcomp2) {
  
  rn <- function(nm) {
    return(paste0(nm, "_2"))
  }
  
  na0 <- function(val) {
    if(is.na(val)) {
      return(0)
    } else {
      return(val)
    }
  }
  
  spcomp2 <- rename_with(spcomp2, rn)
  
  spec1 <- spcomp1 %>%
    select(timechunk, route, region, location.bcr) %>%
    distinct()
  spec2 <- spcomp2 %>%
    select(timechunk_2, route_2, region_2, location.bcr_2) %>%
    distinct()
  
  widecomp <- full_join(spcomp1, spcomp2, by = c("species" = "species_2")) %>%
    group_by_all() %>%
    mutate(relative = na0(relative),
           relative_2 = na0(relative_2)) %>% 
    ungroup()
  
  wideturn <- widecomp %>%
    group_by(species) %>%
    mutate(min_relative = min(relative, relative_2)) %>% 
    dplyr::ungroup() %>%
    dplyr::summarize(compturn = sum(min_relative)) %>%
    bind_cols(spec1) %>%
    bind_cols(spec2)
  
  return(wideturn)
  
}


compare_smooths <- function(smooth1, smooth2) {
  
  rn <- function(nm) {
    return(paste0(nm, "_2"))
  }
  
  na0 <- function(val) {
    if(is.na(val)) {
      return(0)
    } else {
      return(val)
    }
  }
  
  smooth2 <- rename_with(smooth2, rn)
  
  spec1 <- smooth1 %>%
    select(timechunk, route, region, location.bcr) %>%
    distinct()
  spec2 <- smooth2 %>%
    select(timechunk_2, route_2, region_2, location.bcr_2) %>%
    distinct()
  
  widesmooth <- left_join(smooth1, smooth2, by = c("mass" = "mass_2")) 
  
  wideover <- widesmooth %>%
    group_by(mass) %>%
    mutate(min_density = min(density, density_2)) %>% 
    dplyr::ungroup() %>%
    dplyr::summarize(overlap = sum(min_density))%>%
    bind_cols(spec1) %>%
    bind_cols(spec2)
  
  return(wideover)
  
}



compare_pair <- function(all_spcomp, all_smooths, rt1, rg1, bcr1,chunk1, rt2, rg2, bcr2, chunk2) {
  
  spcomp1 <- pull_species_comp(all_spcomp, rt1, rg1, bcr1, chunk1)
  spcomp2 <- pull_species_comp(all_spcomp, rt2, rg2, bcr2, chunk2)
  
  
  smooth1 <- pull_smooths(all_smooths, rt1, rg1, bcr1, chunk1)
  smooth2 <- pull_smooths(all_smooths, rt2, rg2, bcr2, chunk2)
  
  
  species_comparison <- compare_species(spcomp1, spcomp2)
  isd_comparison <- compare_smooths(smooth1, smooth2)
  
  both_comparisons <- left_join(species_comparison, isd_comparison)
  
  return(both_comparisons)
  
}

```


```{r}

# get list of all routes

all_routes <- all_smooths %>% 
  select(route, region, location.bcr) %>%
  distinct() %>%
  mutate(line = row_number())

all_pairs <- expand.grid(r1 = 1:nrow(all_routes), r2 = 1:nrow(all_routes), chunk = c("start", "end")) %>%
  as.data.frame() %>%
  filter(r1 != r2) %>%
  group_by_all() %>%
  mutate(smaller = min(r1, r2),
         larger = max(r1, r2)) %>%
  ungroup() %>%
  select(smaller, larger, chunk) %>%
  distinct() %>%
  left_join(all_routes, by = c("smaller" = "line")) %>%
  left_join(all_routes, by = c("larger" = "line"))

all_comparisons <- list() 

for(i in 1:nrow(all_pairs)) {
  specs <- as.list(all_pairs[i, ])
  
  all_comparisons[[i]] <- compare_pair(all_spcomposition, all_smooths, specs$route.x, specs$region.x, specs$location.bcr.x, specs$chunk, specs$route.y, specs$region.y, specs$location.bcr.y, specs$chunk)
  
}


comparisons <- bind_rows(all_comparisons)

paired_comparisons <- comparisons %>%
  tidyr::pivot_wider(id_cols = c(route, region, location.bcr, route_2, region_2, location.bcr_2), names_from = timechunk, values_from = c(compturn, overlap)) 

ggplot(comparisons, aes(timechunk, overlap)) +
  geom_violin()

```

So I'm not actually seeing a signal of declining beta diversity over this timespan for this bcr. 

Beta diversity obtained as the pairwise overlap in species composition among all possible pairs of routes in the region (relative abundance sp1r1 - sp1r2). 

If routes were becoming more similar in species composition over time, we would expect these pairwise overlap values to be **larger** in the later time period - instead they are pretty much a cloud around the 1:1 line:

```{r}

ggplot(paired_comparisons, aes(compturn_start, compturn_end)) + 
  geom_point() +
  geom_abline(slope = 1, intercept = 0)
```
