---
title: "5 sites"
output:
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(drake)
library(dplyr)
library(ggplot2)
library(ghibli)

theme_set(theme_bw())
## Set up the cache and config
  db <- DBI::dbConnect(RSQLite::SQLite(), here::here("drake-cache-small.sqlite"))
  cache <- storr::storr_dbi("datatable", "keystable", db)
  cache$del(key = "lock", namespace = "session")
  
  loadd(all_overlaps, cache = cache)
loadd(all_smooths, cache = cache)  
loadd(all_svs, cache = cache)
loadd(all_composition, cache = cache)

DBI::dbDisconnect(db)
rm(cache)
rm(db)

```


```{r}

all_overlaps <- left_join(all_overlaps, all_composition)

ggplot(filter(all_overlaps, !is.na(sim_seed)), aes(overlap)) + 
  geom_histogram() +
  facet_wrap(vars(location.routename)) +
  geom_histogram(data = filter(all_overlaps, is.na(sim_seed)), fill = "orange") +
  geom_histogram(data = filter(all_overlaps, is.na(sim_seed)), aes(x = composition_overlap), fill = "green")

```

So a few things here.

1. Observed overlap (orange smidges) isn't generally extraordinary compared to the overlap you get by reassigning species randomly (i.e. without respect to body size).
2. For 3 of these, observed ISD is pretty much on the nose of the overlap in species composition (green smidges). Only for New Hartford and Livermore is there appreciably higher overlap in the ISD than in species composition (indicating any degree of like-for-like substitutions).
3. My intuition gets creeped out when the overlap in the ISD is *less* than the overlap in species composition. I *think* this can occur because the ISD involves intraspecific variation in size, so the ISD can change a little - even if the species composition stays 100% the same - because of size sampling.

