---
title: "Interaction lm thinking"
output: 
    github_document:
      toc: true
      df_print: kable
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(BBSsize)
library(rwar)
theme_set(theme_bw())
```


# Demo dataset

```{r}

d <- BBSsize::hartland

names(d)


begin_years <- c(1994:1998)

end_years <- c(2014:2018)
```

# Simulate ISD

```{r}

d_isd <- BBSsize::simulate_isd_ts(d, isd_seed = 1994)

names(d_isd)

head(d_isd$isd)

```

# Community-level properties

## Compute community-level properties over time

```{r}


d_svs <- get_annual_svs(d_isd$isd)

ggplot(d_svs, aes(year, abundance)) + geom_line()
ggplot(d_svs, aes(year, energy)) + geom_line()
ggplot(d_svs, aes(year, biomass)) + geom_line()

ggplot(d_svs, aes(year, mean_energy)) + geom_line()
ggplot(d_svs, aes(year, mean_biomass)) + geom_line()

```


## Compare how different currencies have changed over time

* Focus on first and last 5 years, because this is the temporal window handled by ISDs and with no gaps

### First and last 5 years

```{r}


d_svs_beginend <- pull_caps(d_svs, begin_years, end_years)


```

#### Raw values and scaling

```{r}

d_svs_raw_long <- d_svs_beginend %>%
  select(-isd_seed) %>%
  tidyr::pivot_longer(-c(timeperiod, year), names_to = "currency", values_to = "val") %>%
  filter(currency %in% c("abundance", "energy", "biomass"))

ggplot(d_svs_raw_long, aes(year, val, color = currency)) + geom_point() + facet_wrap(vars(currency), scales = "free_y") + theme(legend.position = "none")

```

Note the vastly different scales/units of measure for the different currencies. 

Rescale within each currency - there are 2 ways to do this (well, many ways).

`scale(sqrt(value))`: This is what Dornelas 2019 used. It bothers me because you end up with weirdly symmetric statistical properties.

`(val - min(val)) / range(val)`: Rangescaling, results in a number 0-1 inclusive.

### Interaction lm on scale(sqrt) scaling

```{r}

d_svs_ss <- d_svs_beginend %>%
  dplyr::mutate(
    abundance = scale(sqrt(abundance)),
    biomass = scale(sqrt(biomass)),
    energy = scale(sqrt(energy))
  )

d_svs_ss_long <- d_svs_ss %>%
  select(timeperiod, year, abundance, energy, biomass) %>%
  tidyr::pivot_longer(-c(timeperiod, year), names_to = "currency", values_to = "val")

ggplot(d_svs_ss_long, aes(year, val, color = currency)) + geom_point() + facet_wrap(vars(currency), scales = "free_y") + theme(legend.position = "none")

d_svs_ss_ilm <- lm(val ~ timeperiod * currency, data = d_svs_ss_long)

summary(d_svs_ss_ilm)

```

### Interaction lm on rangescale scaling

```{r}

d_svs_rs <- d_svs_beginend %>%
  dplyr::mutate(
    abundance = rangescale(abundance),
    biomass = rangescale(biomass),
    energy = rangescale(energy)
  )

d_svs_rs_long <- d_svs_rs %>%
  select(timeperiod, year, abundance, energy, biomass) %>%
  tidyr::pivot_longer(-c(timeperiod, year), names_to = "currency", values_to = "val")

ggplot(d_svs_rs_long, aes(year, val, color = currency)) + geom_point() + facet_wrap(vars(currency), scales = "free_y") + theme(legend.position = "none")

d_svs_rs_ilm <- lm(val ~ timeperiod * currency, data = d_svs_rs_long)

summary(d_svs_rs_ilm)

```
## Compare ISDs and species composition over time

### ISD over time

Note there is some seed dependence in fitting the gmms, such that overlap can vary.

```{r}
set.seed(1977)
isd_overlap <- compare_isds(d_isd$isd, begin_years, end_years)
isd_overlap
```

### Species composition over time

```{r}
begin_composition <- colSums(d$abundance[1:5,])

end_composition <- colSums(d$abundance[(nrow(d$abundance) - 4): nrow(d$abundance), ])

begin_relabund <- begin_composition / sum(begin_composition)

end_relabund <- end_composition / sum(end_composition)

relabund <- data.frame(
  begin = begin_relabund,
  end = end_relabund,
  beginsp = names(begin_relabund),
  endsp = names(end_relabund)
) 

relabund_change <- relabund %>%
  group_by(beginsp) %>%
  summarize(minRel = min(begin, end)) %>%
  ungroup() %>%
  select(minRel) %>%
  summarize(sp_turnover = 1-sum(minRel))

relabund_change

be_matrix <- bind_rows(begin_composition, end_composition)

be_diss <- vegan::vegdist(be_matrix)

be_diss

compare_species_composition(d)
```


# All variables

```{r}
# 
# community_change <- data.frame(
#   isd_turnover = isd_overlap$isd_turnover[1],
#   species_turnover = relabund_change$sp_turnover[1],
#   bcd = be_diss[[1]])
# 
# outputs <- all_sv_change %>%
#   bind_cols(community_change) %>%
#   bind_cols(d$metadata$location) %>%
#   mutate(beginyears = toString(begin_years),
#          endyears = toString(end_years))
# 
# outputs
# 
# all_core_analyses(d, begin_years, end_years, isd_seed = 1994)
# all_core_analyses(d, isd_seed = 1994)
```

# Variable descriptions

```{r}
# 
# var_names <- data.frame(
#   var_name = colnames(outputs),
#   description = ""
# )
# 
# #write.csv(var_names, "var_descriptions.csv", row.names = F)
# 
# old_var_names <- read.csv("var_descriptions.csv")
# 
# if(any(var_names$var_name != old_var_names$var_name)) {
#   print("vars have changed")
# }
# 
# old_var_names
```
