---
title: "Temporal change"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    github_document:
        toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.dim = c(4,3))
library(dplyr)
library(ggplot2)
theme_set(theme_bw())
library(multipanelfigure)

#allComps <- read.csv(here::here("all_bird_comps_temporal_continental.csv"))

## Set up the cache and config
db <- DBI::dbConnect(RSQLite::SQLite(), here::here("drake-cache-bird-pairs-temporal.sqlite"), synchronous = NULL)
cache <- storr::storr_dbi("datatable", "keystable", db)
cache$del(key = "lock", namespace = "session")
allComps <- drake::readd(allComps, cache = cache)



DBI::dbDisconnect(db)
rm(cache)
print("Completed OK")
obs <- filter(allComps, sim < 0)

```

# 1. Observed dynamics

## State variables (E and N)

```{r, fig.dim = c(4,9)}

quantile(obs$e_logr, probs = c(.025, .5, .975))



mlr_hist <- ggplot(obs, aes(exp(mean_m_logr))) +
  geom_histogram() +
  scale_x_log10()
elr_hist <- ggplot(obs, aes(exp(e_logr))) +
  geom_histogram() +
  scale_x_log10()
nlr_hist <- ggplot(obs, aes(exp(n_logr))) +
  geom_histogram() +
  scale_x_log10()
e_n_l <- ggplot(obs, aes(n_logr, e_logr)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0)

lr_hists <- multi_panel_figure(columns = 1, rows =4) %>%
  fill_panel(mlr_hist) %>%
  fill_panel(elr_hist) %>%
  fill_panel(nlr_hist) %>%
  fill_panel(e_n_l)

lr_hists

summary(lm(e_logr ~ n_logr, obs))
```

* Total E skews weakly towards increasing (median = ```r exp(.2)```, 95% interval ```r exp(c(-.58, .85))```; ```r mean(obs$e_logr > 0)``` increase. 
* Most changes are not dramatic (only 10% of **increases** are more than a doubling; only 3% of **decreases** are less than halving) 
* These figures represent just raw values. You could repeat with a t-test comparing first 5 and last 5 years, or a linear model across the whole timeseries. I lean **against** a hierarchical model, because I am more interested in the heterogeneity of slopes than in an "overall" slope. But hey, you could do it.
* Change in E is strongly predicted by change in N; r2 = .78


## ISD and species compositonal overlap

```{r}

ggplot(obs, aes(isd_overlap)) +
  geom_histogram()

ggplot(obs, aes(species_overlap, isd_overlap)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0)

summary(obs$isd_overlap)

quantile(obs$isd_overlap, probs = c(.025, .5, .975))

summary(lm(isd_overlap ~ species_overlap, obs))

```


* ISDs are highly conserved (median .8, 95% interval .63 - .88)
* Change in the ISD is strongly related to change in species composition (linear model r2 = .66, p << 0.01)
* ISDs are almost always more conserved than species overlap. This is to some extent expected; the ISD can't be (much) less conserved than species overlap. (I'm 99% sure exceptions occur because of intraspecific variation, in this case introduced via sampling).

* I think the increasing E means that there is probably a weak overrepresentation of sites where the mean body size increases over time, but I'm not sure how strong that will be because
    * The slopes are heterogeneous
    * Mass and energy don't scale 1:1, and averaging can be tricky on the intuition
    

```{r}

ggplot(obs, aes(1- species_overlap, abs(mean_m_logr))) + geom_point()

ggplot(obs, aes(1- isd_overlap, (mean_m_logr))) + geom_point()


```


# Residuals

```{r}

e_n_lm <- (lm(e_logr ~ n_logr, obs))

obs_resid <- obs %>%
  mutate(res = resid(e_n_lm))

ggplot(obs_resid, aes(n_logr, e_logr, color = abs(res))) + geom_point()+ scale_color_viridis_c(option = "turbo")


ggplot(obs_resid, aes(1-isd_overlap, abs(res))) + geom_point()

```
