---
title: "How many models come out which way"
output: 
      github_document:
          df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(rwar)
library(BBSsize)
library(brms)
library(tidybayes)
library(ggplot2)
theme_set(theme_bw())

load(here::here("aspirational_structure", "pipelines", "portable_results_core.Rds"))
working_datasets <- read.csv(here::here("aspirational_structure", "supporting_data", "working_routes_max10.csv"))

```


```{r}


all_winners %>%
  group_by(simtype, currency, model) %>%
  summarize(count = dplyr::n())


```



```{r, fig.dim = c(6,9)}

actual_qis <- all_qis %>% 
  filter(simtype == "actual") 
actual_qis_95 <- filter(actual_qis, .width == .95)

ggplot(actual_qis, aes(timeperiodend, matssname)) + geom_pointinterval(aes(xmin = timeperiodend.lower, xmax = timeperiodend.upper, width = .width)) + geom_vline(xintercept = 0) + facet_wrap(vars(currency), scales = "free")

```

Of models of actual,

```{r}

actual_qis_95 %>%
  group_by(currency) %>%
  summarize(percent_with_slope = mean(!is.na(timeperiodend)))

```

Of models with a slope...

```{r}

actual_qis_95 %>%
  group_by(currency) %>%
  summarize(percent_decreasing = mean(timeperiodend.upper < 0, na.rm = T),
            percent_increasing = mean(timeperiodend.lower > 0, na.rm = T),
            n_with_slope = sum(!is.na(timeperiodend)))

```

using a 95% CI above or below 0. This can sum to less than 1 if there is a model with an interaction in which the abundance-slope is over 0, but the currency offset is nonzero. 



Of all models, here's the proportion with an interaction:

```{r}

actual_qis_95 %>%
  group_by(currency) %>%
  summarize(percent_interaction = mean(!is.na(`timeperiodend:sourcecurrency`), na.rm = T))

```

Of models with an interaction...

```{r, fig.dim = c(6,9)}


actual_qis_95 %>%
  group_by(currency) %>%
  summarize(percent_currency_above_abund = mean((`timeperiodend:sourcecurrency.lower` >0), na.rm = T),
            percent_currency_below_abund = mean((`timeperiodend:sourcecurrency.upper` <0), na.rm = T),
            n_with_interaction = sum(!is.na(`timeperiodend:sourcecurrency`)))


ggplot(actual_qis, aes(`timeperiodend:sourcecurrency`, matssname)) + geom_pointinterval(aes(xmin = `timeperiodend:sourcecurrency.lower`, xmax = `timeperiodend:sourcecurrency.upper`, width = .width)) + geom_vline(xintercept = 0) + facet_wrap(vars(currency), scales = "free")

```

```{r}

raw_changes <- all_sims %>%
  filter(simtype == "actual") %>%
  group_by(matssname, source, timeperiod) %>%
  summarize(total_energy = mean(total_energy),
            total_biomass = mean(total_biomass)) %>%
  ungroup() %>%
  tidyr::pivot_wider(id_cols = c(matssname, source), names_from = timeperiod, values_from = c(total_energy, total_biomass)) %>%
  mutate(energy_ratio = total_energy_end / total_energy_begin,
         biomass_ratio = total_biomass_end / total_biomass_begin) %>%
  select(matssname, source, energy_ratio, biomass_ratio) %>%
  tidyr::pivot_wider(id_cols = matssname, names_from = source, values_from = c(energy_ratio, biomass_ratio))

sig_changes <- raw_changes %>%
  left_join(actual_qis_95) %>%
  mutate(abund_slope_nonzero = (timeperiodend.upper < 0 | timeperiodend.lower > 0) & !is.na(timeperiodend),
         currency_offset_nonzero =(`timeperiodend:sourcecurrency.upper` < 0 | `timeperiodend:sourcecurrency.lower` > 0) & !is.na(`timeperiodend:sourcecurrency`))
 
ggplot(sig_changes, aes(energy_ratio_abundance, energy_ratio_currency)) + geom_point(alpha = .1) + geom_point(data =filter(sig_changes, abund_slope_nonzero), aes(color = currency_offset_nonzero), alpha = 1) + geom_abline(slope= 1, intercept = 0) + geom_vline(xintercept = 1) + geom_hline(yintercept = 1) + theme(legend.position = "bottom") + ggtitle("Energy")


ggplot(sig_changes, aes(biomass_ratio_abundance, biomass_ratio_currency)) + geom_point(alpha = .1) + geom_point(data =filter(sig_changes, !is.na(timeperiodend)), aes(color = currency_offset_nonzero), alpha = 1) + geom_abline(slope= 1, intercept = 0) + geom_vline(xintercept = 1) + geom_hline(yintercept = 1) + theme(legend.position = "bottom") + ggtitle("Biomass")



```


## Filtered to max 10

Here is the above, but with a maximum of 10 routes per BCR (n = 238 overall)


```{r}


all_winners_f <- filter(all_winners, matssname %in% working_datasets$matssname)
all_qis_f <- filter(all_qis, matssname %in% working_datasets$matssnam)
all_winners_f %>%
  group_by(simtype, currency, model) %>%
  summarize(count = dplyr::n()) 


```



```{r, fig.dim = c(6,9)}

actual_qis_f <- all_qis_f %>% 
  filter(simtype == "actual") 
actual_qis_95_f <- filter(actual_qis_f, .width == .95)

ggplot(actual_qis_f, aes(timeperiodend, matssname)) + geom_pointinterval(aes(xmin = timeperiodend.lower, xmax = timeperiodend.upper, width = .width)) + geom_vline(xintercept = 0) + facet_wrap(vars(currency), scales = "free")

```

Of models of actual,

```{r}

actual_qis_95_f %>%
  group_by(currency) %>%
  summarize(percent_with_slope = mean(!is.na(timeperiodend)))

```

Of models with a slope...

```{r}

actual_qis_95_f %>%
  group_by(currency) %>%
  summarize(percent_decreasing = mean(timeperiodend.upper < 0, na.rm = T),
            percent_increasing = mean(timeperiodend.lower > 0, na.rm = T),
            n_with_slope = sum(!is.na(timeperiodend)))

```

using a 95% CI above or below 0. This can sum to less than 1 if there is a model with an interaction in which the abundance-slope is over 0, but the currency offset is nonzero. 



Of all models, here's the proportion with an interaction:

```{r}

actual_qis_95_f %>%
  group_by(currency) %>%
  summarize(percent_interaction = mean(!is.na(`timeperiodend:sourcecurrency`), na.rm = T))

```

Of models with an interaction...

```{r, fig.dim = c(6,9)}


actual_qis_95_f %>%
  group_by(currency) %>%
  summarize(percent_currency_above_abund = mean((`timeperiodend:sourcecurrency.lower` >0), na.rm = T),
            percent_currency_below_abund = mean((`timeperiodend:sourcecurrency.upper` <0), na.rm = T),
            n_with_interaction = sum(!is.na(`timeperiodend:sourcecurrency`)))


ggplot(actual_qis_f, aes(`timeperiodend:sourcecurrency`, matssname)) + geom_pointinterval(aes(xmin = `timeperiodend:sourcecurrency.lower`, xmax = `timeperiodend:sourcecurrency.upper`, width = .width)) + geom_vline(xintercept = 0) + facet_wrap(vars(currency), scales = "free")

```
